A properly labelled proof system for intuitionistic modal logics

### 2019-09-18 caleidoscope/caleidoscope.tex
making the completeness proof more beautiful

\begin{definition}
	Let $\M = \langle W, \rel, \le, V \rangle$ be a model and $\B, \Gamma \Rightarrow \Delta$ a sequent. We say that $\M \Vdash \B, \Gamma \Rightarrow \Delta$ if there exists $\f$ such that:
	\begin{enumerate}
		\item For all $\labels{x}{A} \in \Gamma$, we have $\M, f(\lb x) \Vdash \fm A$.
		\item For all $\lb x, \lb y \in \B$, we have $f(\lb x) \rel f(\lb y)$.
		\item For all $\futs xy \in \B$, we have $f(\lb x) \le f(\lb y)$.
	\end{enumerate}
	then for all $\labels{z}{B} \in \Delta$, we have that $\M, f(\lb z) \Vdash \fm B$. We say that $\M \not \Vdash \B, \Gamma \Rightarrow \Delta$ if it is not the case that $\M \Vdash \B, \Gamma \Rightarrow \Delta$.
\end{definition}

@@

\todo{an introduction? do not repeat the theorem?}
\begin{theorem}
	For any formula $A$. If $A$ is a theorem of $\IK$ then $A$ is provable in $\labIKp +\labrn{cut}$.
\end{theorem}

@@

	\kone

@@

	\ktwo

@@

		$\vlderivation {
			\vlin{\rlabrn{\IMP}}
			{}
			{\Rightarrow \labels{x}{\DIA (A \OR B) \IMP (\DIA A \OR \DIA B)}}
			{\vlin {\llabrn{\DIA}}
				{}
				{x \le y, \labels{y}{\DIA (A \OR B)} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}
				{\vliin{\llabrn{\OR}}{}{x \le y, yRz, \labels{z}{A \OR B} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}{\vlhy{\circledast \hspace{40mm}}}{\vlhy{\star}}}}
		}$

@@

	After applying the rule $\llabrn{\OR}$, our proof continues with two premises. On one side, the premise that follows from $\circledast$:

	\bigskip
	\begin{center}
		$\vlderivation{\vlin{\llabrn{\OR}}{}{\circledast}{\vlin {\rlabrn{\OR}}
				{}
				{x \le y, yRz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}
				{\vlin {\rlabrn{\DIA}}
					{}
					{x \le y, yRz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}}
					{\vlin {\rn{refl}}
						{}
						{x \le y, yRz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A}, \labels{z}{A}, \labels{y}{\DIA B}}
						{\vlin {\rn{id}}
							{}
							{x \le y, z \le z, yRz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A}, \labels{z}{A}, \labels{y}{\DIA B}}
							{\vlhy {}}}}}}}$
	\end{center}	
	And on the other hand, we have that the derivation continues from $\star$ like this:

	\begin{center}
		$\vlderivation{\vlin {\llabrn{\OR}}
			{}
			{\star}
			{\vlin {\rlabrn{\OR}}
				{}
				{x \le y, yRz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}
				{\vlin {\rlabrn{\DIA}}
					{}
					{x \le y, yRz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}}
					{\vlin {\rn{refl}}
						{}
						{x \le y, yRz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}, \labels{z}{B}}
						{\vlin {\rn{id}}
							{}
							{x \le y, z \le z, yRz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}, \labels{z}{B}}
							{\vlhy {}}}}}}}$
	\end{center}

>>>>

\begin{definition}
	Let $\M = \langle W, \rel, \le, V \rangle$ be a model and $\B, \Gamma \Rightarrow \Delta$ a sequent. We say that $\M \Vdash \B, \Gamma \Rightarrow \Delta$ if for all $\f$ such that:
	\begin{enumerate}
		\item For all $\labels{x}{A} \in \Gamma$, we have $\M, f(\lb x) \Vdash \fm A$.
		\item For all $\lb x, \lb y \in \B$, we have $f(\lb x) \rel f(\lb y)$.
		\item For all $\futs xy \in \B$, we have $f(\lb x) \le f(\lb y)$.
	\end{enumerate}
	then there exists $\labels{z}{B} \in \Delta$ such that $\M, f(\lb z) \Vdash \fm B$. We say that $\M \not \Vdash \B, \Gamma \Rightarrow \Delta$ if it is not the case that $\M \Vdash \B, \Gamma \Rightarrow \Delta$.
\end{definition}

@@

\todo{an introduction? do not repeat the theorem?}
\begin{theorem}
	For any formula $\fm A$. If $\fm A$ is a theorem of $\IK$ then $\fm A$ is provable in $\labIKp +\labrn{cut}$.
\end{theorem}

@@

		\begin{center}
			$\vlderivation{
				\vlin{\rlabrn{\IMP}}
				{\lb y \mbox{ fresh}}
				{\Rightarrow \labels{x}{\BOX (A \IMP B) \IMP (\BOX A \IMP \BOX B)}}
				{\vlin {\rlabrn{\IMP}}
					{\lb z \mbox{ fresh}}
					{\futs xy, \labels{y}{\BOX(A \IMP B)} \Rightarrow \labels{y}{\BOX A \IMP \BOX B}}
					{\vlin {\rlabrn{\BOX}}
						{\lb w, \lb u \mbox{ fresh}}
						{\futs xy, \futs yz, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A} \Rightarrow \labels{z}{\BOX B}}
						{\vlin {\llabrn{\BOX}}
							{}
							{\futs xy, \futs yz, \futs zw, \accs wu, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A} \Rightarrow \labels{u}{B}}
							{\vlin {\color{red}\rn{trans}}
								{}
								{\futs xy, \futs yz, \futs zw, \accs wu, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A} \Rightarrow \labels{u}{B}}
								{\vlin {\llabrn{\BOX}}
									{}
									{\futs xy, \futs yz, \futs zw, \futs yw, \accs wu, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A} \Rightarrow \labels{u}{B}}
									{\vlin {\color{red}\rn{refl}}
										{}
										{\futs xy, \futs yz, \futs zw, \futs yw, \accs wu, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}, \labels{u}{A \IMP B} \Rightarrow \labels{u}{B}}
										{\vliin{\llabrn{\IMP}}
											{}
											{\futs xy, \futs yz, \futs zw, \futs yw, \futs uu, \accs wu, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}, \labels{u}{A \IMP B} \Rightarrow \labels{u}{B}}
											{\vlin {\rn{id}}
												{}
												{G, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}, \labels{u}{A \IMP B} \Rightarrow \labels{u}{B}, \labels{u}{A}}
												{\vlhy {}}}
											{\vlin {\rn{id}}
												{}
												{G, \labels{y}{\BOX(A \IMP B)}, \labels{z}{\BOX A}, \labels{u}{A}, \labels{u}{A \IMP B}, \labels{u}{B} \Rightarrow \labels{u}{B}}
												{\vlhy {}}}}}}}}}}
			}$
		\end{center}

		where $G$ is equal to: $\futs xy, \futs yz, \futs zw, \futs yw, \futs uu, \accs wu$.

@@

\begin{center}
	$\vlderivation {
		\vlin{\rlabrn{\IMP}}
		{\lb y \mbox{ fresh}}
		{ \Rightarrow \labels{x}{\BOX (A \IMP B) \IMP (\DIA A \IMP \DIA B)}}
		{\vlin {\rlabrn{\IMP}}
			{\lb z \mbox{ fresh}}
			{ \futs xy, \labels{y}{\BOX (A \IMP B)} \Rightarrow \labels{y}{(\DIA A \IMP \DIA B)}}
			{\vlin {\llabrn{\DIA}}
				{\lb u \mbox{ fresh}}
				{\futs xy, \futs yz, \labels{y}{\BOX (A \IMP B)}, \labels{z}{\DIA A} \Rightarrow \labels{z}{\DIA B}}
				{\vlin{\rlabrn{\DIA}}
					{}
					{\futs xy, \futs yz, \accs zu, \labels{y}{\BOX (A \IMP B)}, \labels{u}{A} \Rightarrow \labels{z}{\DIA B}}
					{\vlin {\llabrn{\BOX}}
						{}
						{\futs xy, \futs yz, \accs zu, \labels{y}{\BOX (A \IMP B)}, \labels{u}{A} \Rightarrow \labels{z}{\DIA B}, \labels{u}{B}}
						{\vlin {\color{red}\rn{refl}}
							{}
							{\futs xy, \futs yz, \accs zu, \labels{y}{\BOX (A \IMP B)}, \labels{u}{A}, \labels{u}{A \IMP B} \Rightarrow \labels{z}{\DIA B}, \labels{u}{B}}
							{\vliin{\llabrn{\IMP}}
								{}
								{\futs xy, \futs yz, \accs zu, \futs uu, \labels{y}{\BOX (A \IMP B)}, \labels{u}{A}, \labels{u}{A \IMP B} \Rightarrow \labels{z}{\DIA B}, \labels{u}{B}}
								{\vlin {\rn{id}}
									{}
									{T, \labels{y}{\BOX (A \IMP B)}, \labels{u}{A}, \labels{u}{A \IMP B} \Rightarrow \labels{z}{\DIA B}, \labels{u}{B}, \labels{u}{A}}
									{\vlhy {}}}
								{\vlin {\rn{id}}
									{}
									{T, \labels{y}{\BOX (A \IMP B)}, \labels{u}{A}, \labels{u}{A \IMP B}, \labels{u}{B} \Rightarrow \labels{z}{\DIA B}, \labels{u}{B}}
									{\vlhy {}}}								}}}}}}
	}$
\end{center}

where $T$ is equal to $\futs xy, \futs yz, \accs zu, \futs uu$.

@@

		$\vlderivation {
			\vlin{\rlabrn{\IMP}}
			{}
			{\Rightarrow \labels{x}{\DIA (A \OR B) \IMP (\DIA A \OR \DIA B)}}
			{\vlin {\llabrn{\DIA}}
				{}
				{\futs xy, \labels{y}{\DIA (A \OR B)} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}
				{\vliin{\llabrn{\OR}}{}{\futs xy, \accs yz, \labels{z}{A \OR B} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}{\vlin {\rlabrn{\OR}}
						{}
						{\futs xy, \accs yz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}
						{\vlin {\rlabrn{\DIA}}
							{}
							{\futs xy, \accs yz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}}
							{\vlin {\color{red}{\rn{refl}}}
								{}
								{\futs xy, \accs yz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A}, \labels{z}{A}, \labels{y}{\DIA B}}
								{\vlin {\rn{id}}
									{}
									{\futs xy, \futs zz, \accs yz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A}, \labels{z}{A}, \labels{y}{\DIA B}}
									{\vlhy {}}}}}}{\vlin {\rlabrn{\OR}}
						{}
						{\futs xy, \accs yz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}
						{\vlin {\rlabrn{\DIA}}
							{}
							{\futs xy, \accs yz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}}
							{\vlin {\color{red}{\rn{refl}}}
								{}
								{\futs xy, \accs yz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}, \labels{z}{B}}
								{\vlin {\rn{id}}
									{}
									{\futs xy, \futs zz, \accs yz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}, \labels{z}{B}}
									{\vlhy {}}}}}}}}
		}$

@@
